---
title: "a3_task3_alyssa_kibbe"
author: "Alyssa Kibbe"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```
# Exploring Urchin Dataset
## Preparing Urchins data & exploration
```{r}
urchins <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names() 

urchins <- urchins %>% 
  mutate(date = mdy(date))

class(urchins$date)
  
urchin_summary<-urchins %>%   
  group_by(common_name) %>% 
  summarize(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size),
            n_size= n())
kbl(urchin_summary) %>% 
  kable_styling()

#Percent Difference
(6.776 - 4.470)/((4.470+6.776)/2)*100
#Actual Difference
6.776 - 4.470

ggplot(data = urchins, aes(x = common_name, 
                           y = size))+
  geom_boxplot()+
  labs(x = "Urchin Species", y = "Diameter (cm)")

ggplot(data = urchins, aes(x = size))+
  geom_histogram(bins = 20)+
  facet_wrap(~common_name)+
  labs(x = "Diameter (cm)", y= "Count")

```



#### Exploration results
Data appears to be normally distributed as seen with histogram. Boxplot shows a few values at the high tail. Red Urchins have a a larger mean diameter but a slightly larger standard deviation compared to Purple Urchins. The percent difference between the means is 51.59% while the actual difference is 2.306.

## Purple Urchins
```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(new_date = ymd(date)) %>% 
  mutate(urchin_year = year(new_date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```

### Summary Table & Graph

```{r}
count_purple<- purple_urchins %>%
  count(site_full, urchin_year)

ggplot(data = count_purple, aes(x = urchin_year, y = n)) +
       geom_col(aes(colour = site_full), show.legend = FALSE)+
  facet_wrap(~site_full)+
  theme_bw()+
  labs(x = "Year", y = "Purple Urchin Count")
  
```












